---
title: "Quick EDA on mpg dataset"
author: "Osayame Eghosa"
date: "`r Sys.Date()`"
output:
  pdf_document: default
  html_document: default
  word_document: default
---

```{r echo=T, results='hide', warning=FALSE, error=FALSE, message=FALSE}
#installing needed library and restricting filter to work based on dplyr
library(tidyverse)
library(conflicted)
conflicts_prefer(dplyr::filter)


```

In the code above, we are installing the conflicted package if it is not already installed. Then, we load the tidyverse library, which includes several packages for data manipulation and visualization, and the conflicted library. The conflicts_prefer() function from conflicted ensures that when there is a conflict between function names, we prefer the one from the dplyr package.

```{r}
#Running quick EDA on the dataset
data()
summary(mpg)
glimpse(mpg)

```


Here we perform some quick exploratory data analysis (EDA) on the mpg dataset. We use functions like data() to load the dataset, summary() to get summary statistics, and glimpse() to get a concise summary of the dataset's structure.

```{r}
# Filter data
filter(mpg, cty >=20)
mpg_efficient <- filter(mpg, cty >=20)
mpg_ford <- filter(mpg, manufacturer == "ford")

```

In this section, we use the filter() function from dplyr to subset the data based on specific conditions. We filter the mpg dataset to include only rows where city mileage (cty) is greater than or equal to 20 and create a new dataset mpg_efficient with these filtered rows. We also create another dataset mpg_ford by filtering rows where the manufacturer is "ford".

```{r}
# Adding new column
mpg_metric <- mutate(mpg, cty_metric = 0.425144 * cty)

# Introducing pipe operators
mpg_metric <- mpg %>% 
  mutate(cty_metric = 0.425144 * cty)

mpg %>% 
  group_by(class) %>% 
  summarise(mean(cty), median(cty))

```

Here, we add a new column cty_metric to the mpg dataset by multiplying the cty column by a conversion factor. We demonstrate two ways of achieving this: using the mutate() function directly and using pipe operators (%\>%). Additionally, we group the data by the class variable and calculate the mean and median city mileage for each class.

```{r pressure, echo=TRUE, warning=FALSE, error=FALSE, message=FALSE}
# Data visualization with ggplot
ggplot(mpg, aes(x = cty) )+
  geom_histogram()+
  labs(x = "City Mileage")

ggplot(mpg, aes(x = cty) )+
  geom_freqpoly()+
  labs(x = "City Mileage")

ggplot(mpg, aes(x = cty) )+
  geom_histogram()+
  geom_freqpoly()+
  labs(x = "City Mileage")

ggplot(mpg, aes(x = cty, 
                y = hwy) )+
  geom_point()+
  geom_smooth(method = "lm")

ggplot(mpg, aes(x = cty, 
                y = hwy,
                color = class) )+
  geom_point()+
  scale_color_brewer(palette = "Dark2")

```

Finally, we create several data visualizations using the ggplot2 package. We generate histograms, frequency polygons, scatter plots, and linear regression plots to explore the relationship between variables in the mpg dataset.

